---
title: "Chapter 4 Excercises"
output: github_document
---

Code and solutions for Chapter 4 of the [Statistical Rethinking 2 Ed.](https://xcelab.net/rm/statistical-rethinking/) textbook by R. McElreath.

```{r load-packages}
library(rethinking)
library(tidyverse)
library(ggthemes)
```

Set the palette and the running theme for ggplot2.

```{r set-theme-palette}
theme_set(theme_bw())
theme_update(axis.text.x = element_text(
angle = -45,
hjust = 0,
vjust = 0.5
))
```

## Easy

### 4E1

Line 1

### 4E2

2

### 4E3

$$ Pr(mu, sigma\|h) = Product-i(Normal(h-i\|mu, sigma)\*Normal(mu\| 0, 10)\*Exponential(sigma\| 1)) \\ Integral Integral Product-i(Normal(h-i\|mu, sigma)\*Normal(mu\| 0, 10)\*Exponential(sigma\| 1)) dmu dgamma $$

### 4E4

Line 2

### 4E5

3

## Medium

### 4M1

```{r}
set.seed(42)

n_samples <- 1e4

ex4M1 <- tibble(
  mu = rnorm(n_samples, 0, 10),
  sigma = rexp(n_samples, 1)
) %>%
  mutate(y = rnorm(n_samples, mu, sigma))

glimpse(ex4M1)
```

```{r}
ex4M1 %>%
  ggplot(aes(x = y)) +
  geom_density()
```

### 4M2

```{r}
ex4M2_formula <- alist(
  y ~ dnorm(mu, sigma),
  mu ~ dnorm(0, 10),
  sigma ~ dexp(1)
)
```

### 4M3

y-i \~ Normal(mu-i, sigma)

mu-i = a + b \* w-i

a \~ Normal(0, 10)

b \~ Uniform(0, 1)

sigma \~ Exponential(1)

### 4M4

height-i \~ Normal(mu-i, sigma)

mu-i = a + b \* year-i

a = Normal(178, 20)

b = Log-Normal(0, 1)

sigma = Uniform(0, 50)

### 4M5

Same as 4M4

### 4M6

height-i \~ Normal(mu-i, sigma)

mu-i = a + b \* year-i

a = Normal(178, 20)

b = Log-Normal(0, 1)

sigma = Uniform(0, 10)

### 4M7

```{r}
data(Howell1)
d <- tibble(Howell1)

d2 <- d %>%
  filter(age >= 18) %>%
  mutate(weight_centered = weight - mean(weight))

m4.3 <- quap(
  alist(
    height ~ dnorm(mu, sigma),
    mu <- a + b* weight_centered,
    a ~ dnorm(178, 20),
    b ~ dlnorm(0, 1),
    sigma ~ dunif(0,50)), 
  data = d2)

m4.3bis <- quap(
  alist(
    height ~ dnorm(mu, sigma),
    mu <- a + b* weight,
    a ~ dnorm(178, 20),
    b ~ dlnorm(0, 1),
    sigma ~ dunif(0,50)),
  data = d2)
```

```{r}
precis(m4.3)
```

```{r}
round(vcov(m4.3), 3)
```

```{r}
precis(m4.3bis)
```

```{r}
round(vcov(m4.3bis), 3)
```

```{r}
n_samples <- 1e4
n_grid <- 100

samples_4_3 <- tibble(extract.samples(m4.3, n_samples)) 
samples_4_3b <- tibble(extract.samples(m4.3bis, n_samples))

grid_4_3 <- seq(from = min(d2$weight_centered), 
                to = max(d2$weight_centered), 
                length.out = n_grid)

grid_4_3b <- seq(from = min(d2$weight), 
                 to = max(d2$weight), 
                 length.out = n_grid)

params_mean_4_3 <- samples_4_3 %>%
  summarise(across(everything(), mean))

params_mean_4_3b <- samples_4_3b %>%
  summarise(across(everything(), mean))
```

```{r}
d2 %>%
  ggplot(aes(x = weight_centered, y = height)) +
  geom_point(shape = 21,
             color = "navyblue") +
  geom_abline(data = params_mean_4_3,
              aes(slope = b,
                  intercept = a))
```

```{r}
d2 %>%
  ggplot(aes(x = weight, y = height)) +
  geom_point(shape = 21,
             color = "navyblue") +
  geom_abline(data = params_mean_4_3b,
              aes(slope = b,
                  intercept = a)) +
  coord_cartesian(xlim = c(-1, 70), ylim = c(-1, 185))
```

```{r}
mu_4_3 <- samples_4_3 %>%
  mutate(mu = map2(a, b, ~ .x + .y * grid_4_3),
         weight_centered = list(grid_4_3)) %>%
  unnest(c("mu", "weight_centered")) %>%
  group_by(weight_centered) %>%
  summarise(mu_mean = mean(mu),
            mu_HPDI_89 = list(HPDI(mu))) %>%
  unnest_wider(mu_HPDI_89)

mu_4_3
```

```{r}
mu_4_3 %>%
  ggplot(aes(x = weight_centered,
             y = mu_mean)) +
  geom_line(color = "black") +
  geom_ribbon(aes(ymin = `|0.89`,
                  ymax = `0.89|`),
              fill = "grey70",
              alpha = 0.3) +
  geom_point(data = d2,
             aes(x = weight_centered,
                 y = height),
             color = "navyblue",
             shape = 21) +
  labs(x = "Weight - Mean Weight",
       y = "Height")
```

```{r}
mu_4_3b <- samples_4_3b %>%
  mutate(mu = map2(a, b, ~ .x + .y * grid_4_3b),
         weight = list(grid_4_3b)) %>%
  unnest(c("mu", "weight")) %>%
  group_by(weight) %>%
  summarise(mu_mean = mean(mu),
            mu_HPDI_89 = list(HPDI(mu))) %>%
  unnest_wider(mu_HPDI_89)

mu_4_3b
```

```{r}
mu_4_3b %>%
  ggplot(aes(x = weight,
             y = mu_mean)) +
  geom_line(color = "black") +
  geom_ribbon(aes(ymin = `|0.89`,
                  ymax = `0.89|`),
              fill = "grey70",
              alpha = 0.3) +
  geom_point(data = d2,
             aes(x = weight,
                y = height),
             color = "navyblue",
             shape = 21) +
  labs(x = "Weight",
       y = "Height")
```

### 4M8

```{r}
data("cherry_blossoms")
d <- cherry_blossoms
precis(d)
```

```{r 4.73}
d2 <- d[complete.cases(d$doy),]
num_knots <- 15
knot_list <- quantile(d2$year, 
                      probs = seq(0, 1, length.out = num_knots))
```

```{r 4.74}
library(splines)
B <- bs(d2$year,
        knots = knot_list[-c(1, num_knots)],
        degree = 3,
        intercept = TRUE)

str(B)
```

```{r}
m4.7 <- quap(
  alist(
    D ~ dnorm(mu, sigma),
    mu <- a + B %*% w,
    a ~ dnorm(100, 10),
    w ~ dnorm(0,2),
    sigma ~ dexp(1)
  ), 
  data = list(D = d2$doy, B = B),
  start = list(w = rep(0, ncol(B)))
)
```

```{r}
post <- extract.samples(m4.7)
precis(post, depth = 2)
```

```{r}
str(post)
```

```{r}
w <- apply(post$w, 2, mean)
plot(NULL, xlim = range(d2$year), ylim = c(-6,6),
     xlab = "year", ylab = "basis * weight")
for(i in 1:ncol(B)){
  lines(d2$year, w[i]*B[,i])
}
```

```{r}
mu <- link(m4.7)
mu_PI <- apply(mu, 2, PI, 0.97)

plot(d2$year, 
     d2$doy, 
     col = col.alpha(rangi2, 0.3),
     pch = 16)
shade(mu_PI, 
      d2$year,
      col = col.alpha("black", 0.5))
```

## Hard

Document the information about the analysis session

```{r sessionInfo, include=TRUE, echo=TRUE, results='markup'}
sessionInfo()
```
