---
title: "Chapter 5 Excercises"
output: github_document
---

Code and solutions for Chapter 5 of the [Statistical Rethinking 2 Ed.](https://xcelab.net/rm/statistical-rethinking/) textbook by R. McElreath.

```{r load-packages}
library(rethinking)
library(tidyverse)
library(ggthemes)
```

Set the palette and the running theme for ggplot2.

```{r set-theme-palette}
theme_set(theme_bw())
theme_update(axis.text.x = element_text(
angle = -45,
hjust = 0,
vjust = 0.5
))
```

## Easy

### 5E1

(4) mu[i] = A + Bx * x[i] + Bz * z[i]

### 5E2

L[i] ~ Normal(mu[i], sigma)

mu[i] = A + Bd * animal_diversity[i] + Bd * plant_diversity[i]

### 5E3

T[i] ~ Normal(mu[i], sigma)

mu[i] = A + Bf * funding[i] + Bs * size[i]

Both Bf and Bd should have a positive value

### 5E4

(4) mu[i] = Aa * a[i] + Ab * b[i] + Ac * c[i] + Ad * d[i]
(5) mu[i] = Aa * (1 - b[i] - c[i] - d[i]) + Ab * b[i] + Ac * c[i] + Ad * d[i]

In (4) the coefficients represent mean values for the different levels of the variable: Aa mean value when the variable is A, Ab mean value when the variable is B, Ac mean value when the variable is C, Ad mean value when the variable is D.

In (5):

* If the variable = A, then mu[i] = Aa
* If the variable = B, then mu[i] = Aa - Aa + Ab, or mu[i] = Ab
* If the variable = B, then mu[i] = Aa - Aa + Ac, or mu[i] = Ac
* If the variable = B, then mu[i] = Aa - Aa + Ad, or mu[i] = Ad

These are the same coefficients seen in (4).

## Medium

### 5M1

Predictor 1: Maximum wind speed on any given day between October and April for any Italian city above 100k inhabitants

Predictor 2: Is the city Trieste?

Outcome: Normalized number of buffet restaurants specialized in boiled pork and sauerkraut for any Italian city above 100k inhabitants

Max wind speed in winter <- Triestiness -> Number of buffet restaurants 

### 5M2

Predictor1: Number of hours worked

Predictor2: Fatigue level

Outcome: Work Output 

Total Work Output  <- (+) Number of hours worked (+) -> Fatigue level (-) > Total Work Output 

## Hard

### 5H1

```{r}
library(rethinking)
library(dagitty)

DMA_dag_3 <- dagitty('dag{M -> A -> D}')
coordinates(DMA_dag_3) <- list(x = c(M = 0, A = 0.5, D = 1), y = c(M = 0, A = 0, D = 0))
drawdag(DMA_dag_3)
```

```{r}
impliedConditionalIndependencies(DMA_dag_3)
```

Yes, the data is consistent with DAG # 3. A, M and D are all correlated in the absence of conditioning on A, and D is independent of M, conditional on A. 

### 5H2

```{r}
library(tidyverse)

data("WaffleDivorce")
d <- tibble(WaffleDivorce)

d <- d %>%
  select(A = MedianAgeMarriage, 
         D = Divorce, 
         M = Marriage) %>%
  transmute(across(everything(), ~ standardize(.x)))

glimpse(d)
```

```{r}
m5H2 <- quap(
  alist(
    ## M -> A -> D
    D ~ dnorm(mu, sigma),
    mu <- a + bM * M + bA * A,
    a ~ dnorm(0, 0.2),
    bM ~ dnorm(0, 0.5),
    bA ~ dnorm(0, 0.5),
    sigma ~ dexp(1),
    ## M -> A
    A ~ dnorm(muA, sigmaA),
    muA <- aMA + bMA * M,
    aMA ~ dnorm(0, 0.2),
    bMA ~ dnorm(0, 0.5),
    sigmaA ~ dexp(1)
  ), data = d
)

precis(m5H2)
```

```{r}
set.seed(2971)

sim_dat <- tibble(M = seq(-2, 2, length.out = 30))

m5H2_sim <- sim(m5H2, data = sim_dat, vars = c("A", "D"))
```

```{r}
plot(sim_dat$M, colMeans(m5H2_sim$D), ylim = c(-2, 2), type = "l")
shade(apply(m5H2_sim$D, 2, PI), sim_dat$M)
```
```{r}
plot(sim_dat$M, colMeans(m5H2_sim$A), ylim = c(-2, 2), type = "l")
shade(apply(m5H2_sim$A, 2, PI), sim_dat$M)
```



Document the information about the analysis session

```{r sessionInfo, include=TRUE, echo=TRUE, results='markup'}
sessionInfo()
```
